Thermal performance of fish is explained by an interplay between physiology, behaviour, and ecology

### 2018-07-06 MS_AMNat.tex
AmNat submission

\documentclass[11pt]{article}
\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
\usepackage{gensymb}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[authoryear,sectionbib,sort]{natbib}
\linespread{1.7}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{titlesec}
\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}

@@

\date{}

\begin{document}

@@

Temperature, through its effects on individual physiology, is a dominant driver of species ecology and bio-geography (\citealt{deutsch_climate_2015,pinsky_marine_2013}). As a consequence, current and predicted temperature increases under climate change will act as a strong agent of change in many ecosystems (\citealt{deutsch_climate_2015, stuart-smith_thermal_2015, parmesan_globally_2003, walther_ecological_2002}). However, the nature of these changes can be difficult to predict as temperature effects scale from individuals to species and ecosystems. Through this cascade of scales, incorrect or approximate model assumptions at the individual scale can have disproportionate effects on ecosystem level outcomes (\citealt{brander_overconfidence_2013, lefevre_models_2017}). In marine fish, for example, recent predictions of decreasing organism size, and resulting decreases in the size of fisheries catches (\citealt{cheung_shrinking_2013}), have been criticized as overly simplistic and not in line with physiological constraints (\citealt{brander_overconfidence_2013, lefevre_models_2017,jutfelt_oxygen-and_2018}).

Although there are conceptual and model frameworks to explain aspects of thermal performance and ecological responses to temperature (\citealt{fry_effects_1947,brown_toward_2004,pauly_sound_2017,portner_oxygen-and_2010}), many of these remain controversial as they appear limited in their generality or predictive capacity (\citealt{lefevre_models_2017,jutfelt_oxygen-and_2018}). To our knowledge, no general theoretical framework exists to quantitatively explain and predict changes in  ecological rates, such as observed change in growth and asymptotic size \cite[e.g., the temerature-size rule in ectotherms;][]{angilletta_temperature_2004,atkinson_temperature_1994}), or attack rates (\citealt{englund_temperature_2011,rall_universal_2012}) with temperature, from fundamental physiological processes. Instead, and as advocated by the metabolic theory of ecology (\citealt{brown_toward_2004}), ecological theory often treats ecological rates as being directly temperature dependent, without a direct link to the underlying physiological drivers (\cite[e.g., ][]{angilletta_temperature_2004,vucic-pestic_warming_2011, guiet_effects_2016}).

A phenomenological description that assumes a general ecological temperature response often fails to explain heterogeneity in ecological responses (\citealt{angilletta_temperature_2004,rall_universal_2012}). Although ecological rates seem to follow some general patterns in the response to temperature, there is also significant heterogeneity between species and trait groups (\cite[e.g., ][]{angilletta_temperature_2004,englund_temperature_2011,rall_universal_2012}). This leads to difficulties with extrapolation across species or other model components (\cite[e.g., fish in marine ecosystem models; ][]{guiet_effects_2016}). In this case, a deeper understanding of the underlying drivers of thermal responses may be necessary in order to derive general predictions about ecological responses to changing temperatures (\cite[e.g., ][]{vucic-pestic_warming_2011,lefevre_models_2017}). Since the primary effect of temperature of organisms is on individual physiology, a general model to explain ecological response should be grounded in physiology.

Physiologically, a long held view has been that temperature is a controlling factor while oxygen supply sets the physiological limits (\citealt{fry_effects_1947,claireaux_linking_2007,lefevre_are_2016}). How exactly temperature influences ectotherm physiological rates and limits, however, has been a matter of debate, not least because of the variable responses observed among different species. In most species, the standard metabolism (SM; the metabolic cost of maintenance and routine activity such as ventilation) increases near exponentially with temperature. A prevalent view is that the maximum metabolic rate (MMR; the metabolic rate at maximum sustained exercise) has a dome-shaped response to temperature, whereby it can be increased (passively and actively) up to a point, but plateaus or decreases thereafter (\citealt{fry_effects_1947,claireaux_linking_2007,lefevre_are_2016,portner_physiology_2008}). This leads to the view of a uni-modal curve for metabolic scope (MMR minus SM; the available oxygen/energy for additional activity), and suggests that towards the upper end of this curve, organisms will, simply put, run out of oxygen.

This view was encapsulated in the Theory of Oxygen and Capacity Limitation of Temperature (\citealt{portner_oxygen-and_2010}), which suggests that the decrease in metabolic scope towards extreme temperatures limits species ability to sustain core functions such as foraging and growth (i.e., functions beyond SM). In some species, however, maximum metabolic rate increases steadily (\citealt{lefevre_are_2016, verberk_does_2016}), suggesting that oxygen may not be the limiting factor at high temperatures. Indeed, it has been argued that oxygen is unlikely to determine performance for most species over most of their temperature range as oxygen limits are rarely reached during normal activity (\citealt{jutfelt_oxygen-and_2018, holt_climate_2015}).

@@

Our model assumes that physiology is described by two key budgets: the energy and oxygen budgets (\citealt{holt_climate_2015,holt_climate_2014}). We assume that animals will adapt activity levels to optimise available energy for growth and reproduction relative to mortality risk. Available energy is determined either by food capture, food processing capacity, or by available oxygen. We further assume that temperature acts directly on rates that are determined by enzymatic activity: digestive activity (via maximum consumption) and metabolic costs. Consequently, temperature only acts on ecological rates (e.g., actual feeding rates) via optimisation of activity levels.

\subsection*{Model description}
Ectotherms adjust the relative amounts of time ($\tau$) spent on metabolically costly activity and resting/hiding to optimise the net energy gain relative to mortality (\citealt{gilliam_habitat_1987}). In the following, we refer to $\tau$ as the activity fraction for sake of generality.   Since both energy gain and metabolic losses are sensitive to temperature and oxygen limitations, both the activity level and the net energy gain will be subject to these environmental constraints. Their interplay thus determines available energy for growth and reproduction.

@@

Metabolic demands ($D(w,T)$) are standard metabolism ($\propto k w^n$), which scales with exponent $n<1$, and active metabolism $\propto \tau k_a w$, which scales proportional to mass owing to muscular demands scaling approximately isometrically with weight (\citealt{glazier_activity_2009,brett1965relation}, and the activity fraction. Temperature scaling of metabolic rates (standard and active metabolism, and maximum consumption rate) is determined by enzymatic processes (e.g., digestion, glycolysis \citealt{jeschke_predator_2002, sentis_parsing_2013}) and approimated by an Arrhenious scaling $c(T) = e^{E_a(T-T_0)/(bTT_0)}$, where $E_a$ is the activation energy, assumed constant, $T_0$ is the reference temperature (here 15 degrees Celsius), and $b$ is the Bolzman constant. Note that we only scale rates related to enzymatic activity with temperature, we do assume that ecological rates such as foraging rates or activity are a direct function of temperature. Rather, they are modulated by an individual's behavioural response to temperature driven physiological changes.

The oxygen budget $P_{0_2}(w,T)$ (or aerobic scope) follows a similar form to the mass budget:
\begin{align}
  P_{0_2}(w,T) &= S_{O_2}(w,T) - D_{O_2}(w,T) \\
        &= S_{O_2}c_{O_2}(T) w^n - \omega c(T) \left( \beta f(w,T) h w^q + k w^n + k_a w \right).
\end{align}
Demand ($D_{O_2}(w,T)$) is the sum of oxygen used for all metabolic processes in Eq.~\ref{eq:P} (except assimilation losses), with $\omega$ being amount of oxygen required per mass.  The oxygen supply ($S_{O_2}c_{O_2}(T)$) scales with body weight as $w^n$ multiplied by a flexible dome-shaped function that can emulate both a dome-shaped maximum oxygen supply (MOS) as well as a MOS that increase continuously up to a lethal temperature (fig./~\ref{fig:O2_fig_used}). The maximum oxygen consumption is the oxygen consumption during maximal activity level that can be sustained over some time, and is usually termed the maximum metabolic rate (MMR). Although the MMR is often used synonymously with both maximum oxygen supply and demand, in some species the maximal oxygen consumption ($D^{max}_{O_2}(T)$) is not reached at maximum activity levels, but rather during digestion (\citealt{priede_metabolic_1985}). We use MMR and oxygen supply interchangeably, although maximum oxygen demand (i.e., at full activity) may be lower than the MMR. We assume that oxygen supply is temperature dependent and follows a flexible dome-shaped function (\citealt{gnauck2013freshwater, lefrancois_influence_2003}):

@@

In pour model, fish will adjust their activity fraction $\tau$ to maximise fitness. We use Gilliam's rule as a fitness proxy \citealt{gilliam_habitat_1987}:
\begin{equation}
  \tau^* = \text{argmax}_{\tau} \{ \frac{P(w,T)}{M(w)} \}.
\end{equation}
This optimisation represents a "short-sighted" fitness optimisation that does not account for future changes in conditions, and is appropriate for optimisation in stable environments (\citealt{sainmont_effective_2015}).  Mortality scales with activity fraction and weight as $w^{q-1}$ (\citealt{andersen_how_2009, hartvig_food_2011}):
\begin{equation}
  \label{eq:M}
  M(w) = (\rho+\eta\tau) w^{q-1},
\end{equation}
$\rho$ is mortality at mass $w=1$ and $\tau=0$, that is, with no activity beyond that covered by standard metabolism.  Activity is limited by available oxygen, such that the aerobic scope $P_{O_2}$ is not allowed to be negative over the time scales considered. In other words, we consider time scales that are long enough to ignore the ability of many ectotherms to go into oxygen debt, or to switch to anaerobic metabolism for limited periods of time. This means we assume that animals will adjust their foraging effort to optimise fitness given temperature and oxygen constraints. Note, that this optimal foraging assumption drives ecological responses as a consequence of physiological constraints, rather than as a direct response to temperature itself.

@@

Performance itself is a vague concept that is often used without definition in the relevant literature, but it is only defined in the context of a particular physiological or demographic parameter, with potentially different thermal response curves (\citealt{jutfelt_oxygen-and_2018}). For instance, even though growth in ectotherms is often impacted by temperature (\citealt{angilletta_temperature_2004}), measured performance may depend strongly on what aspect of growth is measured --- whether it is the growth increment at a particular size, growth efficiency per unit intake, the attained asymptotic size, or parameters of a fitted growth curve. We use and compare four performance measures with broad ecological and practical implications:

@@

The slope of maturation reaction norms is evolutionarily determined by the strength of the co-variation between growth and mortality for a given population in a given environment. Strongly positive co-variation leads to the relatively flat reaction norms observed for most fish populations (\citealt{marty_impact_2011}). This covariation is probably the consequence of good growth conditions (e.g., from increased prey density) altering the baseline mortality $\rho$ and the risk of foraging $\eta$ (e.g., by attracting predators). We did not explicitly model these interactions here (aside from the dependence of mortality on $\tau$), but rather assumed that reaction norms evolved over regimes of relatively stable temperature and growth variation in the past. Consequently, we assumed that the evolved slope of the reaction norm is a fixed trait over the time-scales considered here for a particular species or population, and for simplicity and generality we assume a flat reaction norm (i.e., allocation to reproduction is a function of size only; fig./~\ref{fig:MRN}), although sloping reaction norms can be formulated and used in our framework (see appendix \ref{sec:MRN}). The intercept of this reaction norm was found numerically by maximising fitness ($R_0$, see below) at the reference temperature for our simulations (15\degree C), and this intercept was assumed fixed as temperatures change.

The allocation was parametrised as 
\[ \phi(w^*) = 1/(1+\exp(-cw^*))\],
where $w^*$ is the intercept of the reaction norm, and $c$ determines how rapidly energy allocation shifts from somatic growth to reproduction.

@@

\[R_0(T) = \int_0^\infty \phi(w(t),w^*)P(w(t),T)S_{0 \rightarrow t}(T) \text{d}t, \text{, where} \\
S_{0 \rightarrow t}(T) = \int_0^t \exp(-M(w(t),T))\text{d}t \]

<< setup,echo=FALSE,include=F>>=
knitr::opts_chunk$set(echo = F, message = F,cache=T,autodep = T)

require(dplyr)
require(tidyr)
require(rlang)
require(cowplot)

source('model.R')
cols <- c('#4477AA','#44AA77','#AA7744','#AA4488','#DDDD77')

@

<<Scenarios>>=

int_m <- 500
int_t <- 500
int_temp <- 30

# Sluggish small
ss <- list(c=0.3,
           beta=0.1,
           delta=4,
           Ea=0.52, 
           gamma=60, 
           h=30, 
           k=1,
           lO=0.5, 
           m=10, 
           n=0.88, 
           o=1,
           O2crit=2, 
           omega=0.25, 
           p=0.8,
           P50=4.2, 
           phi=0.25, 
           q=0.8,
           shape=3,
           temp=seq(5,26,l=int_temp),
           Topt=20, 
           temp_ref=15,
           slope=0,
           tmax=20,
           tr=0.5,
           M=0.1,
           v=6,
           lm=int_m,
           nu=0.8-1,
           dt=int_t,
           min_m=-2,
           max_m= 3
)

# fast small0.8-1
fs <- ss
fs$lO <- 1
fs$k <- 1.5
fs$h <- 60
fs$v <- 1
fs$M <- 1
fs$nu<- 0.8-1
fs$n <- 0.75
fs$delta <- 4
fs$o <- 1
fs$p <- 0.8
fs$q <- 0.8
fs$tmax <- 10

fl <- fs
fl$shape <- 0.1
fl$Topt <- 25

sl <- ss
sl$shape <- 0.1
sl$Topt <- 25

ffl <- fs
ffl$gamma <- 20

fsl <- ss
fsl$gamma <- 20

ffh <- fs
ffl$gamma <- 80

fsh <- ss
fsl$gamma <- 80

scs <- list(ss,fs,sl,fl)
fscs <- list(fsh,ffh,fsl,ffl)

baseline <- bind_cols(data_frame(Scenario = as.factor(rep(c('M strategy','P strategy','M strategy','P strategy'),each=int_temp)),
                                  Food = as.factor(rep(c('High','Low'),each=int_temp*2))),
                       bind_rows(lapply(fscs, get_model_out,seq(0.05,1,l=int_temp))))

scenarios <- bind_cols(data_frame(Scenario = as.factor(rep(c('M strategy','P strategy','M strategy','P strategy'),each=int_temp)),
                                  O2lim = as.factor(rep(c('Domed','Rising'),each=int_temp*2))),
                       bind_rows(lapply(scs, plot_data_temp)))

scn <- rep(unique(scenarios$Scenario),2)
scz <- rep(unique(scenarios$O2lim),each=2)

@

<<scenarios,fig.cap='Available energy $P$ (Blue solid line), efficiency (green dashed line), and the ratio of $P$ to $M$ (orange dotted line) are impacted as a function of changing activity at constant temperature ($T=15$\\degree C) for a growing fish ($L_{\\infty}\\sim 30$cm) at 10cm length (10g). Responses are shown for slow- (a) and fast strategy (b). All rates are plotted relative to their maximum for each trait scenario, the optimal activity level $\\tau^*$ is indicated by the dotted vertical line.',fig.align='center',fig.height=3,fig.width=6>>=

lab <- as_labeller(c(
                   `M strategy Low`  ='a',
                   `P strategy Low` = 'b'))

sc_bl <- baseline  %>%
  filter(Food == 'Low') %>% 
  group_by(Scenario) %>%
  mutate(Consumption=sc(Consumption),
         Efficiency=sc(Efficiency),
         Feeding=sc(`Feeding level`),
         `Available Energy` = ifelse(`C for growth`<0,0,`C for growth`),
         `Relative M` = sc(M),
         `Energy` = sc(`Available Energy`),
         Fitness = sc(`Available Energy`/M)) %>% 
  tidyr::gather(Rate,value,Energy,Efficiency,`Relative M`,Fitness) %>%
  mutate(Rate = factor(Rate,levels = c('Energy','Efficiency','Fitness')),
         Dispatch = paste(Scenario,Food))

tstar <- baseline  %>%
  filter(Food == 'Low') %>%
  group_by(Scenario, Food) %>%
  mutate(`Available Energy` = ifelse(`C for growth`<0,0,`C for growth`),
         Fitness = sc(`Available Energy`/M)) %>% 
  summarise(tstar = tau[which.max(Fitness)]) %>%
  mutate(Dispatch = paste(Scenario,Food))

ggplot() + 
  geom_line(aes(x=tau, y=value, linetype=Rate,col=Rate,group=Rate),data=sc_bl) +
  geom_vline(aes(xintercept=tstar),linetype=3,data=tstar) +
  theme_cowplot()+
  facet_wrap(~Dispatch,labeller = lab,ncol=2)+
  theme(strip.text = element_text(hjust = 0.05),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank())+
  scale_linetype_manual(values=c(1:4,1),guide='none')+
  scale_color_manual(values = cols,guide='none')+
   labs(x = expression('Activity ' (tau)),
       y='Relative rate')

@

@@

To implement this axis, we used the result that species with a more
active, production oriented life-history (e.g., predatory pelagic
fish) have a higher standard metabolism and lower weight scaling of
metabolic costs
(\citealt{priede_metabolic_1985,killen_intraspecific_2010}). We assumed
that higher standard metabolism is due to increased digestive capacity
(i.e., is used for gut maintenance), though high muscle mass and a
larger heart will also contribute to higher standard metabolism in
active species (\citealt{priede_metabolic_1985}). In practice, we assumed
that approximately 50\% of the standard metabolic cost is due to gut
maintenance, such that a doubling of the maximum ingestion leads to a
50\% increase in standard metabolic cost. We further assumed that such active species have a less effective refuge from predators and therefore have a higher constant mortality, but lower mortality related to activity (i.e., $M_{s}(w) = (0.1+6\tau) w^{q-1}$ and $M_{f}(w) = (1+\tau) w^{q-1}$). Exact parameter values for these trait scenarios are given in Table \ref{tab:parameters}. Together, these assumed trait differences lead to very different ecological and bio-energetic responses of slow- and fast strategists (fig./~\ref{fig:scenarios}), with $\tau_{\text{opt}}$ found at lower activity levels for slow-strategists, as high activity induces exceedingly high mortality and decreasing energy efficiency (i.e., available energy relative to food intake) at high activity levels. A slower increase in available energy and M with $\tau$ for fast strategists leads to a higher $\tau_{\text{opt}}$.

>>>>

\documentclass[11pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
\usepackage{gensymb}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[authoryear,sectionbib,sort]{natbib}
\linespread{1.7}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{titlesec}
\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}

@@

\date{}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

Temperature, through its effects on individual physiology, is a dominant driver of species ecology and bio-geography (\cite{deutsch_climate_2015,pinsky_marine_2013}). As a consequence, current and predicted temperature increases under climate change will act as a strong agent of change in many ecosystems (\cite{deutsch_climate_2015, stuart-smith_thermal_2015, parmesan_globally_2003, walther_ecological_2002}). However, the nature of these changes can be difficult to predict as temperature effects scale from individuals to species and ecosystems. Through this cascade of scales, incorrect or approximate model assumptions at the individual scale can have disproportionate effects on ecosystem level outcomes (\cite{brander_overconfidence_2013, lefevre_models_2017}). In marine fish, for example, recent predictions of decreasing organism size, and resulting decreases in the size of fisheries catches (\cite{cheung_shrinking_2013}), have been criticized as overly simplistic and not in line with physiological constraints (\cite{brander_overconfidence_2013, lefevre_models_2017,jutfelt_oxygen-and_2018}).

Although there are conceptual and model frameworks to explain aspects of thermal performance and ecological responses to temperature (\cite{fry_effects_1947,brown_toward_2004,pauly_sound_2017,portner_oxygen-and_2010}), many of these remain controversial as they appear limited in their generality or predictive capacity (\cite{lefevre_models_2017,jutfelt_oxygen-and_2018}). To our knowledge, no general theoretical framework exists to quantitatively explain and predict changes in  ecological rates, such as observed change in growth and asymptotic size \cite[e.g., the temerature-size rule in ectotherms;][]{angilletta_temperature_2004,atkinson_temperature_1994}), or attack rates (\cite{englund_temperature_2011,rall_universal_2012}) with temperature, from fundamental physiological processes. Instead, and as advocated by the metabolic theory of ecology (\cite{brown_toward_2004}), ecological theory often treats ecological rates as being directly temperature dependent, without a direct link to the underlying physiological drivers (\cite[e.g., ][]{angilletta_temperature_2004,vucic-pestic_warming_2011, guiet_effects_2016}).

A phenomenological description that assumes a general ecological temperature response often fails to explain heterogeneity in ecological responses (\cite{angilletta_temperature_2004,rall_universal_2012}). Although ecological rates seem to follow some general patterns in the response to temperature, there is also significant heterogeneity between species and trait groups (\cite[e.g., ][]{angilletta_temperature_2004,englund_temperature_2011,rall_universal_2012}). This leads to difficulties with extrapolation across species or other model components (\cite[e.g., fish in marine ecosystem models; ][]{guiet_effects_2016}). In this case, a deeper understanding of the underlying drivers of thermal responses may be necessary in order to derive general predictions about ecological responses to changing temperatures (\cite[e.g., ][]{vucic-pestic_warming_2011,lefevre_models_2017}). Since the primary effect of temperature of organisms is on individual physiology, a general model to explain ecological response should be grounded in physiology.

Physiologically, a long held view has been that temperature is a controlling factor while oxygen supply sets the physiological limits (\cite{fry_effects_1947,claireaux_linking_2007,lefevre_are_2016}). How exactly temperature influences ectotherm physiological rates and limits, however, has been a matter of debate, not least because of the variable responses observed among different species. In most species, the standard metabolism (SM; the metabolic cost of maintenance and routine activity such as ventilation) increases near exponentially with temperature. A prevalent view is that the maximum metabolic rate (MMR; the metabolic rate at maximum sustained exercise) has a dome-shaped response to temperature, whereby it can be increased (passively and actively) up to a point, but plateaus or decreases thereafter (\cite{fry_effects_1947,claireaux_linking_2007,lefevre_are_2016,portner_physiology_2008}). This leads to the view of a uni-modal curve for metabolic scope (MMR minus SM; the available oxygen/energy for additional activity), and suggests that towards the upper end of this curve, organisms will, simply put, run out of oxygen.

This view was encapsulated in the Theory of Oxygen and Capacity Limitation of Temperature (\cite{portner_oxygen-and_2010}), which suggests that the decrease in metabolic scope towards extreme temperatures limits species ability to sustain core functions such as foraging and growth (i.e., functions beyond SM). In some species, however, maximum metabolic rate increases steadily (\cite{lefevre_are_2016, verberk_does_2016}), suggesting that oxygen may not be the limiting factor at high temperatures. Indeed, it has been argued that oxygen is unlikely to determine performance for most species over most of their temperature range as oxygen limits are rarely reached during normal activity (\cite{jutfelt_oxygen-and_2018, holt_climate_2015}).

@@

Our model assumes that physiology is described by two key budgets: the energy and oxygen budgets (\cite{holt_climate_2015,holt_climate_2014}). We assume that animals will adapt activity levels to optimise available energy for growth and reproduction relative to mortality risk. Available energy is determined either by food capture, food processing capacity, or by available oxygen. We further assume that temperature acts directly on rates that are determined by enzymatic activity: digestive activity (via maximum consumption) and metabolic costs. Consequently, temperature only acts on ecological rates (e.g., actual feeding rates) via optimisation of activity levels.

\subsection*{Model description}
Ectotherms adjust the relative amounts of time ($\tau$) spent on metabolically costly activity and resting/hiding to optimise the net energy gain relative to mortality (\cite{gilliam_habitat_1987}). In the following, we refer to $\tau$ as the activity fraction for sake of generality.   Since both energy gain and metabolic losses are sensitive to temperature and oxygen limitations, both the activity level and the net energy gain will be subject to these environmental constraints. Their interplay thus determines available energy for growth and reproduction.

@@

Metabolic demands ($D(w,T)$) are standard metabolism ($\propto k w^n$), which scales with exponent $n<1$, and active metabolism $\propto \tau k_a w$, which scales proportional to mass owing to muscular demands scaling approximately isometrically with weight (\cite{glazier_activity_2009,brett1965relation}, and the activity fraction. Temperature scaling of metabolic rates (standard and active metabolism, and maximum consumption rate) is determined by enzymatic processes (e.g., digestion, glycolysis \cite{jeschke_predator_2002, sentis_parsing_2013}) and approimated by an Arrhenious scaling $c(T) = e^{E_a(T-T_0)/(bTT_0)}$, where $E_a$ is the activation energy, assumed constant, $T_0$ is the reference temperature (here 15 degrees Celsius), and $b$ is the Bolzman constant. Note that we only scale rates related to enzymatic activity with temperature, we do assume that ecological rates such as foraging rates or activity are a direct function of temperature. Rather, they are modulated by an individual's behavioural response to temperature driven physiological changes.

The oxygen budget $P_{0_2}(w,T)$ (or aerobic scope) follows a similar form to the mass budget:
\begin{align}
  P_{0_2}(w,T) &= S_{O_2}(w,T) - D_{O_2}(w,T) \\
        &= S_{O_2}c_{O_2}(T) w^n - \omega c(T) \left( \beta f(w,T) h w^q + k w^n + k_a w \right).
\end{align}
Demand ($D_{O_2}(w,T)$) is the sum of oxygen used for all metabolic processes in Eq.~\ref{eq:P} (except assimilation losses), with $\omega$ being amount of oxygen required per mass.  The oxygen supply ($S_{O_2}c_{O_2}(T)$) scales with body weight as $w^n$ multiplied by a flexible dome-shaped function that can emulate both a dome-shaped maximum oxygen supply (MOS) as well as a MOS that increase continuously up to a lethal temperature (fig./~\ref{fig:O2_fig_used}). The maximum oxygen consumption is the oxygen consumption during maximal activity level that can be sustained over some time, and is usually termed the maximum metabolic rate (MMR). Although the MMR is often used synonymously with both maximum oxygen supply and demand, in some species the maximal oxygen consumption ($D^{max}_{O_2}(T)$) is not reached at maximum activity levels, but rather during digestion (\cite{priede_metabolic_1985}). We use MMR and oxygen supply interchangeably, although maximum oxygen demand (i.e., at full activity) may be lower than the MMR. We assume that oxygen supply is temperature dependent and follows a flexible dome-shaped function (\cite{gnauck2013freshwater, lefrancois_influence_2003}):

@@

In pour model, fish will adjust their activity fraction $\tau$ to maximise fitness. We use Gilliam's rule as a fitness proxy \cite{gilliam_habitat_1987}:
\begin{equation}
  \tau^* = \text{argmax}_{\tau} \{ \frac{P(w,T)}{M(w)} \}.
\end{equation}
This optimisation represents a "short-sighted" fitness optimisation that does not account for future changes in conditions, and is appropriate for optimisation in stable environments (\cite{sainmont_effective_2015}).  Mortality scales with activity fraction and weight as $w^{q-1}$ (\cite{andersen_how_2009, hartvig_food_2011}):
\begin{equation}
  \label{eq:M}
  M(w) = (\rho+\eta\tau) w^{q-1},
\end{equation}
$\rho$ is mortality at mass $w=1$ and $\tau=0$, that is, with no activity beyond that covered by standard metabolism.  Activity is limited by available oxygen, such that the aerobic scope $P_{O_2}$ is not allowed to be negative over the time scales considered. In other words, we consider time scales that are long enough to ignore the ability of many ectotherms to go into oxygen debt, or to switch to anaerobic metabolism for limited periods of time. This means we assume that animals will adjust their foraging effort to optimise fitness given temperature and oxygen constraints. Note, that this optimal foraging assumption drives ecological responses as a consequence of physiological constraints, rather than as a direct response to temperature itself.

@@

Performance itself is a vague concept that is often used without definition in the relevant literature, but it is only defined in the context of a particular physiological or demographic parameter, with potentially different thermal response curves (\cite{jutfelt_oxygen-and_2018}). For instance, even though growth in ectotherms is often impacted by temperature (\cite{angilletta_temperature_2004}), measured performance may depend strongly on what aspect of growth is measured --- whether it is the growth increment at a particular size, growth efficiency per unit intake, the attained asymptotic size, or parameters of a fitted growth curve. We use and compare four performance measures with broad ecological and practical implications:

@@

The slope of maturation reaction norms is evolutionarily determined by the strength of the co-variation between growth and mortality for a given population in a given environment. Strongly positive co-variation leads to the relatively flat reaction norms observed for most fish populations (\cite{marty_impact_2011}). This covariation is probably the consequence of good growth conditions (e.g., from increased prey density) altering the baseline mortality $\rho$ and the risk of foraging $\eta$ (e.g., by attracting predators). We did not explicitly model these interactions here (aside from the dependence of mortality on $\tau$), but rather assumed that reaction norms evolved over regimes of relatively stable temperature and growth variation in the past. Consequently, we assumed that the evolved slope of the reaction norm is a fixed trait over the time-scales considered here for a particular species or population, and for simplicity and generality we assume a flat reaction norm (i.e., allocation to reproduction is a function of size only; fig./~\ref{fig:MRN}), although sloping reaction norms can be formulated and used in our framework (see appendix \ref{sec:MRN}). The intercept of this reaction norm was found numerically by maximising fitness ($R_0$, see below) at the reference temperature for our simulations (15\degree C), and this intercept was assumed fixed as temperatures change.

The allocation was parametrised as 
$$\phi(w^*) = 1/(1+\exp(-cw^*))$$,
where $w^*$ is the intercept of the reaction norm, and $c$ determines how rapidly energy allocation shifts from somatic growth to reproduction.

@@

$$R_0(T) = \int_0^\infty \phi(w(t),w^*)P(w(t),T)S_{0 \rightarrow t}(T) \text{d}t, \text{, where}$$
$$S_{0 \rightarrow t}(T) = \int_0^t \exp(-M(w(t),T))\text{d}t$$

\begin{figure}

{\centering \includegraphics[width=\maxwidth]{images/scenarios-1}

}

\caption[Available energy $P$ (Blue solid line), efficiency (green dashed line), and the ratio of $P$ to $M$ (orange dotted line) are impacted as a function of changing activity at constant temperature ($T=15$\degree C) for a growing fish ($L_{\infty}\sim 30$cm) at 10cm length (10g)]{Available energy $P$ (Blue solid line), efficiency (green dashed line), and the ratio of $P$ to $M$ (orange dotted line) are impacted as a function of changing activity at constant temperature ($T=15$\degree C) for a growing fish ($L_{\infty}\sim 30$cm) at 10cm length (10g). Responses are shown for slow- (a) and fast strategy (b). All rates are plotted relative to their maximum for each trait scenario, the optimal activity level $\tau^*$ is indicated by the dotted vertical line.}\label{fig:scenarios}
\end{figure}

@@

To implement this axis, we used the result that species with a more
active, production oriented life-history (e.g., predatory pelagic
fish) have a higher standard metabolism and lower weight scaling of
metabolic costs
(\cite{priede_metabolic_1985,killen_intraspecific_2010}). We assumed
that higher standard metabolism is due to increased digestive capacity
(i.e., is used for gut maintenance), though high muscle mass and a
larger heart will also contribute to higher standard metabolism in
active species (\cite{priede_metabolic_1985}). In practice, we assumed
that approximately 50\% of the standard metabolic cost is due to gut
maintenance, such that a doubling of the maximum ingestion leads to a
50\% increase in standard metabolic cost. We further assumed that such active species have a less effective refuge from predators and therefore have a higher constant mortality, but lower mortality related to activity (i.e., $M_{s}(w) = (0.1+6\tau) w^{q-1}$ and $M_{f}(w) = (1+\tau) w^{q-1}$). Exact parameter values for these trait scenarios are given in Table \ref{tab:parameters}. Together, these assumed trait differences lead to very different ecological and bio-energetic responses of slow- and fast strategists (fig./~\ref{fig:scenarios}), with $\tau_{\text{opt}}$ found at lower activity levels for slow-strategists, as high activity induces exceedingly high mortality and decreasing energy efficiency (i.e., available energy relative to food intake) at high activity levels. A slower increase in available energy and M with $\tau$ for fast strategists leads to a higher $\tau_{\text{opt}}$.
